---
title: "Fractional Differencing Post"
author: "Peter Caya"
date: "2022-09-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Steps: [1] Load bank data into memory for 90 DPD on residential
mortgages. [2] Pick aggregation groups [3] Determine the transformations
to be used for independent variables. • Distributed lag transformation •
Fractional differencing ̉• Level variables. [4] Form the transformations
for the variables. [5] Fit list of models [6] Examine relative
performance of models over time periods; Include discussion of the
recessionary periods.

Define functions:

```{r}
#################################################################################################### 
# Function Name: centerData 
# Description: Function to center the values of a data-series
# Inputs 
# [1] dataValues
# Outputs:
# [1] centeredData 
centerData <- function(dataValues) { 
   dataMin <- min(dataValues, na.rm =T)
   dataMax <- max(dataValues, na.rm =T)
   centeredData <- (dataValues-dataMin)/(dataMax - dataMin)
   return(centeredData) } 
#####################################################################################################
```

# Load Data

```{r}
library(pacman)
p_load(data.table,magrittr,ggplot2)
formattingDetails <- function(){return( theme_bw()+theme(axis.text = element_text(size = 13),
                                                         legend.position = "bottom",
                                                         axis.title = element_text(size = 16 ),
                                                         text = element_text(family = "sans"), 
                                                         title = element_text(size = 20),
                                                         legend.text = element_text(size = 13),
                                                         legend.title = element_blank())      )}

stateFileData <- list.files("Call_Report_App/ID_files/",pattern = "RDS$",full.names = T)
stateFileData <- lapply(X = stateFileData,FUN = function(x){as.data.table(readRDS(x))} ) %>% rbindlist
macroData <- readRDS("C:/Users/peter/Dropbox/Projects/FFIEC/Call_Report_App/Macroeconomic/data_for_all_variables.rds")

pdData <- stateFileData[(Field == "90 days past due")|Field == "Total loans" , ][Type == "Credit Card"  ][`Financial Institution State`!=0][,c("Date","Financial Institution Name","Financial Institution State","Field","value"),with = F]
pdData <- dcast(data = pdData,formula = Date + `Financial Institution Name` + `Financial Institution State`~Field,fun.aggregate = function(x){sum(x,na.rm = T)},value.var = "value") %>% as.data.table

dataforTrainingNational <- pdData[,lapply(X = .SD, FUN = function(x){sum(x,na.rm = T)}),.SDcols = c("90 days past due","Total loans"),by = c("Date" ) ]
dataforTrainingNational <- dataforTraining[,"90+ DPD":= `90 days past due`/`Total loans`]

```

## Plot the results:

```{r}
ggplot(data = dataforTrainingNational) + geom_line(aes(x=Date,`90+ DPD`)) + formattingDetails() + ggtitle("National 90+ DPD")
```

## Normalize the result and plot alongside untransformed variables

```{r}
# Collect macro-economic data
macroVars <- readRDS("[2] Macroeconomic data/untransformedMacro.RDS")
names(macroVars) <- c("Date","Consumer Debt Service Payments","Consumer Price Index", "HPI","10 Year Treasury Yields","GDP","SP 500","Unemployment Rate")
macroVars <- as.data.table(macroVars)
```

```{r}
centeredData <- macroVars[,lapply(X = .SD,function(x){centerData(x)}  ),.SDcols = c( "Consumer Debt Service Payments","Consumer Price Index", "HPI","10 Year Treasury Yields","GDP","SP 500","Unemployment Rate")]
centeredData$Date <- macroVars$Date
centeredData
```

```{r}
dataforTrainingNational$Date <- dataforTrainingNational$Date +1
results <- merge(x =dataforTrainingNational[,c("Date","90+ DPD"),with = F] , y=centeredData , by = "Date") 
centeredData <- results[,lapply(X = .SD,function(x){centerData(x)}  ),.SDcols = c( "90+ DPD","Consumer Debt Service Payments","Consumer Price Index", "HPI","10 Year Treasury Yields","GDP","SP 500","Unemployment Rate")]
centeredData$Date <- results$Date

```

```{r}
centeredDataforPlot <- melt(centeredData,id.vars = "Date")
ggplot(centeredDataforPlot)+geom_line(aes(x = Date , y = value , color = variable)) + formattingDetails()+  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7"))
```

# Variable investigation

## Transform macro-economic variables

Three sets of transformations are used:

1.  Relative relative difference.
2.  The factors and transformations described in Breeden.
3.  Fractional differencing transformations

```{r}
source("C:/Users/peter/Dropbox/Projects/FFIEC/functions for blog post.R")
transformedData <- apply(X = centeredData[,-1],2,
                         FUN = function(x){
                           identifyDifferencingNew(x,dLow = 0,dHigh = 2,size = 10,steps = 20,outputTransformed = TRUE    )}     )



```

```{r}




# For each variable use the following fractional differencing:
# d values from 0 (no change) to 2
# All values will be considered. Nonstationarity and noncorrelation will be prioritized in model selection.








```

## Form distributed lag models

## Form models with fractional differencing
